\Huge Conferences and Cost of Carbon

### 2019-06-03 main.tex
CP

This leads to a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  This seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.
%
Fortunately, there is a well-established mechanism for comparing apples and
oranges---indeed, for comparing almost anything: the magic of markets.  By
mapping choices to currency amounts, markets facilitate individual valuation
of costs and benefits across vastly different activities.\bcp{Give a simple,
  concrete example.} For this reason, many policy experts advocate some form
of carbon pricing, which places a monetary cost on greenhouse gas emissions,
giving a market signal that changes aggregate behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}
Exactly how this should be done is a challenging question, but it has a
clear success 
criterion: The cost should be set so that the world arrives at zero
Emissions within the timeframe that climate scientists tell us is safe.
%
For ACM, this means that the carbon cost of conferences should be
reflected in their monetary cost.

% Expecting that this is coming, what should we rationally do now to ensure
% that ACM, and computer-science research broadly, continue to thrive in a
% world where air travel is more expensive?  Believing that it needs to come
% quickly, what should we ethically do now to accelerate the process?
% Recognizing that the climate change is the defining crisis of our time, how
% can ACM and its members participate in addressing it?

In the longer term, this cost should be imposed from outside, by governments.
Indeed, many countries and states (e.g., California) have already adopted
carbon pricing policies; more will follow.
%
Moreover, although the current market price of carbon is still quite low
(around \$15 / ton), experts agree that this will need to rise significantly
over time to achieve the required effect.
%

%
In the short term, we can choose to do it voluntarily using carbon
offsets.  
\begin{itemize}
\item (What they are.)
\item giving ourselves a ``price signal''
is an effective way to begin raising awareness and changing organizational
behavior.
\item Lots of organizations (universities, companies) are doing this; some
ACM conferences are starting to experiment with voluntary carbon offset fees
at registration time. 
\item There are also some significant
concerns.
\item Concern:
Organizations are not set up to do it.  Our experience with getting
ACM to introduce an offsetting policy shows that it requires significant
effort.
\item Concern: It is hard to choose the price (right now there is an oversupply), and
setting it too low leads to complacency --- it doesn't really bite, so it
doesn't change people's behavior
\item This issue leads many people to argue that offsets "let people off the hook" and encourage
inaction, but we believe it is consistent with encouraging action -- again,
if offsets are priced right.
\end{itemize}

First, we can begin gathering and analyzing data, to better understand the
carbon footprint of our conferences and the likely effects (both
environmental and scientific) of potential changes in the way they are
organized.  \bcp{Put in the most interesting discussion and perhaps show an
  interesting diagram.}

% \bcp{Add some meat to that.
%   \begin{itemize}
%   \item List a few things we can do
%   \begin{itemize}
%   \item journalize conferences
%   \item go virtual
%   \item potential parts of a solution: regional conferences, virtual PC
%   meetings, virtual participation, mega-conferences
%   \end{itemize}
%   \item cite our report
%   \end{itemize}
% }

Ultimately, responding to the inevitability of expensive carbon may well
involve rethinking the whole basis of how computer science is done.
\bcp{Add a bit of meat about how the current economic model may be
  unsustainable and reworking it will involve substantial organizational
  changes.} 
\begin{itemize}
\item monitoring and mitigating carbon costs should be an ACM-level activity
\end{itemize}

>>>>

This leads to a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  This seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.
%
Fortunately, there is a well-established mechanism for comparing apples and
oranges---indeed, for comparing almost anything: the magic of markets.  By
mapping choices to currency amounts, markets facilitate individual valuation
of costs and benefits across vastly different activities.\bcp{Give a simple,
  concrete example.} For this reason, many policy experts advocate some form
of carbon pricing, which places a monetary cost on greenhouse gas emissions,
giving a market signal that changes aggregate behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}
Exactly how this should be done is a challenging question, but it has a
clear success 
criterion: The cost should be set so that the world arrives at zero
Emissions within the timeframe that climate scientists tell us is safe.

In the long term, this cost must be imposed from outside, by governments.
Indeed, many countries and states (e.g., California) have already adopted
carbon pricing policies; more will follow.
%
Moreover, although the current market price of carbon is still quite low
(around \$15 / ton), experts agree that this will need to rise significantly
over time to achieve the required effect.  Bottom line: carbon is going to
be expensive.

Expecting that this is coming, what should we rationally do now to ensure
that ACM, and computer-science research broadly, continue to thrive in a
world where air travel is more expensive?  Believing that it needs to come
quickly, what should we ethically do now to accelerate the process?

A short-term response to both questions: ACM can choose to voluntarily impose
a price on its emissions using carbon offsets. 
\begin{itemize}
\item (What they are.)
\item giving ourselves a ``price signal''
is an effective way to begin raising awareness and changing organizational
behavior.
\item Lots of organizations (universities, companies) are doing this; some
ACM conferences are starting to experiment with voluntary carbon offset fees
at registration time. 
\item There are also some significant
concerns.
\item Concern:
Organizations are not set up to do it.  Our experience with getting
ACM to introduce an offsetting policy shows that it requires significant
effort.
\item Concern: It is hard to choose the price (right now there is an oversupply), and
setting it too low leads to complacency --- it doesn't really bite, so it
doesn't change people's behavior
\item This issue leads many people to argue that offsets "let people off the hook" and encourage
inaction, but we believe it is consistent with encouraging action -- again,
if offsets are priced right.
\end{itemize}

Whether we choose to impose carbon offsets on ourselves or wait for
governments to do it, we need to begin gathering and analyzing data, to
better understand the 
carbon footprint of our conferences and the likely effects (both
environmental and scientific) of potential changes in the way they are
organized.  \bcp{Put in the most interesting discussion and perhaps show an
  interesting diagram.}
\bcp{Add some meat to that.
  \begin{itemize}
  \item List a few things we can do (and whose likely effects we'd need data
  to estimate)
  \begin{itemize}
  \item journalize conferences
  \item go virtual
  \item potential parts of a solution: regional conferences, virtual PC
  meetings, virtual participation, mega-conferences
  \end{itemize}
  \item cite our report
  \end{itemize}
}

Ultimately, responding to the inevitability of expensive carbon may involve
rethinking the whole basis of how academic computer science is done.
\bcp{
\begin{itemize}
\iten Add a bit of meat about how the current economic model may be
  unsustainable and reworking it will involve substantial organizational
  changes.
\item monitoring and mitigating carbon costs should be an ACM-level activity
\end{itemize}
}
